\documentclass[11pt]{article}

% ==================== packages ====================
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{xcolor}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{authblk}
\usepackage{caption}
\captionsetup{font=small}

\newcommand{\raw}{\textsc{raw}}
\renewcommand{\cal}{\textsc{cal}}
\newcommand{\best}[1]{\textbf{#1}}
\newcommand{\vt}{v_t}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator*{\argmin}{arg\,min}

\title{\bfseries Action-Conditional Flow-Matching World Models for\\
Regime-Switching Markets: Strictly-Proper Path-Distribution\\
Training and Few-Step Distillation}

\author{FinFlow Project}
\affil{Course: Principles and Methods of Deep-Learning-based Content Generation}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We study generative \emph{world models} for synthetic financial markets, where the
generator must be (i) \emph{controllable} (conditioned on a latent market regime
that an external agent can set) and (ii) \emph{self-consistent} under
\emph{free-running} multi-step simulation rather than one-step teacher forcing.
Our testbed is a three-regime Markov-switching Heston market with a Monte-Carlo
(MC) option-pricing oracle, which lets us measure both distributional realism
(stylized facts, signature-Wasserstein distance) and downstream financial
fidelity (European-option pricing error against the oracle). We build a two-stage,
action-conditional \emph{Flow-Matching} (FM) world model (variance stage followed
by return stage) and then make two contributions on top of it. First, we propose
a \emph{strictly-proper path-distribution fine-tuning} procedure that replaces the
adversarial Wasserstein-GAN critic of Quant-GAN with a combination of a
signature-kernel maximum-mean-discrepancy (sig-MMD), a Sig-W$_1$ expected-signature
term, and an energy score, all evaluated on a \emph{differentiable free-running
rollout}. This non-adversarial objective monotonically improves calibrated pricing
error from $0.583$ (base teacher) through $0.420$ (adversarial critic) to
$\best{0.180}$. Second, we present a systematic study of one-step and few-step
\emph{distillation} (consistency, Mean-Flow, and Lagrangian flow-map
self-distillation) of the world model, and find that consistency distillation is
the only family that transfers under free rollout, while Mean-Flow and flow-map
students degrade or diverge. Across all $25{+}$ models we surface a robust
\emph{raw$\leftrightarrow$calibrated trade-off}: the base FM teacher is the raw
self-consistency sweet spot ($0.475$, near the $0.165$ MC finite-sample floor),
scheduled sampling gives the best stylized-fact balance (kurtosis $4.41$ vs.\ the
real $4.60$), and aggressive calibration can drive pricing error \emph{below} the
oracle floor --- which we identify and flag as a calibration \emph{artifact}
rather than genuine generative quality. We release all training, distillation and
evaluation code.
\end{abstract}

% =====================================================================
\section{Introduction}
\label{sec:intro}

Generative models of financial time series are used to stress-test trading
strategies, augment scarce data, and price derivatives under scenarios that are
under-represented in history. Unlike images or text, a financial generator is
useful only if it reproduces a battery of \emph{stylized facts}---heavy-tailed
returns, volatility clustering, the leverage effect---and, crucially, if the
\emph{prices of derivatives} computed on its sample paths match those of the true
data-generating process \citep{cont2001empirical}. These two requirements are not
the same: a model can match marginal histograms yet misprice options, or vice
versa.

We take a \emph{world-model} view \citep{ha2018world,alonso2024diamond}: the
generator is an action-conditional simulator that an external agent drives by
choosing, at each step, a discrete market \emph{regime} (normal / high-volatility
/ crash). This framing has two consequences that drive the entire paper. First,
the model is evaluated under \emph{free-running rollout}: we autoregressively
unroll it for a full trading year (252 steps) and never feed back ground-truth
state (no teacher forcing). This is far harder than one-step prediction and
exposes \emph{exposure bias}---the train/test mismatch between teacher-forced
training and free generation \citep{bengio2015scheduled}. Second, controllability
means the architecture must accept an action channel and remain stable when the
agent forces rare regimes.

We instantiate the world model with \emph{Flow Matching}
\citep{lipman2023flow,liu2023rectified,albergo2023stochastic}, a simulation-free
continuous-time generative framework that is more stable to train than diffusion
or GANs and admits principled few-step distillation. Our generator is two-stage
and transition-based: a \emph{variance} flow models the next instantaneous
variance and a \emph{return} flow models the next log-return, each conditioned on
the current state and the regime action. On top of this teacher we make two
contributions.

\paragraph{Contribution 1: strictly-proper path-distribution fine-tuning.}
Quant-GAN \citep{wiese2020quant}, the standard deep baseline for financial series,
trains a generator against a Wasserstein-GAN critic
\citep{gulrajani2017improved}. Adversarial training is unstable and provides no
distributional guarantee. We instead fine-tune the FM world model on a
\emph{differentiable free-running rollout} using a sum of \emph{strictly-proper}
\citep{gneiting2007strictly} path-distribution losses: a signature-kernel MMD
\citep{salvi2021signature,issa2023nonadversarial,gretton2012kernel}, a Sig-W$_1$
expected-signature term \citep{ni2021sigwgan,ni2020conditional}, and an energy
score \citep{pacchiardi2024probabilistic}, together with low-order moment / tail
regularizers. Replacing the adversarial critic with these proper scoring rules
improves calibrated option-pricing RMSE monotonically (\S\ref{sec:results}) and is
the methodological core of the paper.

\paragraph{Contribution 2: a study of few-step world-model distillation.}
A simulator is only practical if a long rollout is cheap. We distill the
multi-step FM teacher into one-/few-step students with three modern recipes:
consistency distillation \citep{song2023consistency,song2024improved}, Mean-Flow
\citep{geng2025meanflow}, and Lagrangian flow-map self-distillation
\citep{boffi2025flowmaps}. Under the demanding free-rollout metric we find that
consistency distillation is the \emph{only} family that preserves financial
fidelity, while Mean-Flow and flow-map students degrade or blow up the tails---a
practically important negative result for one-step generation of long
autoregressive rollouts.

\paragraph{Contribution 3: an honest evaluation protocol and the
raw$\leftrightarrow$calibrated trade-off.} We price options on generated paths
against an independent 100k-path MC oracle and report results both \emph{raw} and
\emph{moment-calibrated}. The real test set itself prices at $0.165$ against the
oracle (a finite-sample floor). We show that a single affine moment correction can
push reported pricing error \emph{below} this floor (e.g.\ one of our own
consistency students reaches $0.118$ while its raw error is $2.33$ and its
kurtosis collapses to $2.85$), and we argue---and demonstrate across models---that
any sub-floor calibrated number is a calibration \emph{artifact}, not generative
quality. Disentangling these two regimes is, we believe, a contribution in its own
right for the financial-generation literature.

Our findings: (i) the base FM teacher is the \emph{raw} self-consistency champion
($0.475$); (ii) scheduled sampling yields the best stylized-fact balance
(kurtosis $4.41$, closest to the real $4.60$); (iii) our strictly-proper
fine-tuning yields the best \emph{honest} calibrated pricing ($0.180$); and (iv)
there is a systematic tension between raw self-consistency and calibrated pricing
that no post-hoc procedure removes.

% =====================================================================
\section{Related Work}
\label{sec:related}

\paragraph{Generative models for financial time series.}
Quant-GAN \citep{wiese2020quant} couples a temporal convolutional generator with a
Lambert-W heavy-tail output transform \citep{goerg2015lambert} and a WGAN-GP
objective. Sig-WGAN and its conditional variant
\citep{ni2021sigwgan,ni2020conditional} replace the learned critic with a
signature-based Wasserstein-1 distance, exploiting the fact that the signature
\citep{lyons2007differential} linearizes path functionals. TimeGAN
\citep{yoon2019timegan} combines adversarial and supervised losses in a learned
embedding. Classical econometric baselines---GARCH and its asymmetric GJR variant
with Student-$t$ innovations \citep{bollerslev1986garch,glosten1993relation,
bollerslev1987conditionally}---remain strong on volatility clustering and are the
reference any deep model must beat.

\paragraph{Flow Matching and world models.}
Flow Matching \citep{lipman2023flow}, rectified flow \citep{liu2023rectified} and
stochastic interpolants \citep{albergo2023stochastic} learn a continuous-time
velocity field that transports noise to data, sidestepping the score-matching and
sampling difficulties of diffusion \citep{ho2020denoising,karras2022elucidating}.
World models \citep{ha2018world,alonso2024diamond} learn action-conditional
simulators of an environment; our work imports this view into quantitative
finance, treating the market regime as the action.

\paragraph{Strictly-proper scoring rules.}
A scoring rule is strictly proper if its expectation is uniquely minimized by the
true distribution \citep{gneiting2007strictly}. The energy score and
kernel/MMD scores \citep{gretton2012kernel} are strictly proper and have been used
to train probabilistic forecasters \citep{pacchiardi2024probabilistic} and Neural
SDEs without a discriminator \citep{issa2023nonadversarial}. We assemble several
such path-level scores into a single non-adversarial fine-tuning objective for an
FM world model.

\paragraph{Few-step distillation.}
Consistency models \citep{song2023consistency,song2024improved} learn a one-step
map that is invariant along the probability-flow ODE. Mean-Flow
\citep{geng2025meanflow} learns the time-averaged velocity for one-step sampling.
Flow-map self-distillation \citep{boffi2025flowmaps} learns a two-time transport
map $X(x,s,t)$ via a Lagrangian objective, enabling a tunable
quality/compute trade-off by composing the map over $K$ sub-intervals. We are, to
our knowledge, the first to benchmark all three under a free-running financial
rollout.

% =====================================================================
\section{Problem Setup}
\label{sec:setup}

\paragraph{Data: three-regime Markov-switching Heston.}
The ground-truth market follows a Heston stochastic-volatility model
\citep{heston1993} whose parameters switch between three latent regimes according
to a Markov chain. Under regime $a\in\{\text{normal},\text{high-vol},\text{crash}\}$,
\begin{align}
dS_t &= \mu_a S_t\,dt + \sqrt{\vt}\,S_t\,dW^S_t, \\
d\vt &= \kappa_a(\theta_a-\vt)\,dt + \xi_a\sqrt{\vt}\,dW^v_t,
\qquad d\langle W^S,W^v\rangle_t = \rho_a\,dt,
\end{align}
with $(\kappa,\theta,\xi,\rho,\mu)$ equal to $(2,0.04,0.3,-0.7,0.05)$,
$(3,0.09,0.5,-0.7,0.05)$, $(4,0.16,0.8,-0.7,-0.2)$ for the three regimes. The
regime evolves by a row-stochastic transition matrix with persistent diagonals
($0.992,0.95,0.90$). We simulate $S_0{=}100$, $v_0{=}0.04$, $dt{=}1/252$, and
$T{=}252$ steps (one trading year). We generate $50$k train, $5$k validation and
$10$k test paths (seed $20260530$), and an independent $100$k-path
\emph{MC oracle} for pricing.

\paragraph{World-model factorization.}
We model the one-step transition autoregressively. With $r_t$ the log-return and
$a_t$ the regime action, the world model factorizes each step as
\begin{equation}
p_\theta(\vt[t{+}1], r_t \mid \vt, r_{t-1}, a_t)
= \underbrace{p^{\mathrm{vol}}_\theta(\vt[t{+}1]\mid \vt, r_{t-1}, a_t)}_{\text{variance stage}}\;
  \underbrace{p^{\mathrm{ret}}_\theta(r_t\mid \vt[t{+}1], \vt, r_{t-1}, a_t)}_{\text{return stage}} ,
\end{equation}
with $r_{-1}{=}0$. Each stage is a conditional Flow-Matching model
(\S\ref{sec:method}). A full simulation is produced by \emph{free-running rollout}:
sample $\vt[t+1]$, then $r_t$, append, and repeat for $252$ steps under an
externally chosen regime sequence.

\paragraph{Evaluation protocol.}
All numbers are computed on a free-running rollout of $5000$ paths
$\times$ $252$ steps with the FM ODE integrated in $16$ steps (\emph{no} teacher
forcing). We report three families of metrics:
\begin{itemize}\itemsep2pt
\item \textbf{Option-pricing RMSE} against the MC oracle: European-call prices over
moneyness $\{0.85,0.90,0.95,1.00,1.05\}$ and maturities $\{0.25,0.5,1.0\}$,
RMSE of (generated vs.\ oracle) prices. This is our primary financial-fidelity
metric. The real test set prices at $\mathbf{0.165}$ against the oracle---a
\emph{finite-sample floor}; no generator should honestly beat it.
\item \textbf{Stylized facts}, chiefly excess kurtosis of pooled log-returns
(real value $\mathbf{4.60}$), plus return / absolute-return autocorrelation and the
leverage correlation.
\item \textbf{Signature-Wasserstein distance} (depth 3) between real and generated
path ensembles.
\end{itemize}

\paragraph{\raw{} vs.\ \cal{}.}
We report pricing RMSE in two modes. \raw{} uses the generator output directly.
\cal{} applies a single affine \emph{moment calibration}: pooled generated returns
are standardized and re-scaled to the data's return mean and standard deviation
before pricing (mirroring the moment correction used by Quant-GAN). Calibration is
legitimate for a practitioner who only needs prices, but it injects exactly two
data-derived numbers (the first two pooled moments) that dominate option prices.
Consequently, \textbf{a \cal{} value below the $0.165$ floor is a calibration
artifact}, not evidence of a better generator. We flag all such rows.

% =====================================================================
\section{Method}
\label{sec:method}

\subsection{Two-stage Flow-Matching world model (the teacher)}
Flow Matching \citep{lipman2023flow} learns a velocity field $v_\theta(x,\tau,c)$
that transports a Gaussian prior at flow-time $\tau{=}0$ to the data at $\tau{=}1$,
conditioned on context $c$. With the linear (rectified) interpolant
$x_\tau=(1-\tau)\,x_0+\tau\,x_1$, $x_0\sim\mathcal N(0,I)$, the training objective is
\begin{equation}
\mathcal L_{\mathrm{FM}}
= \E_{\tau\sim U[0,1],\,x_1\sim\mathcal D,\,x_0}
\big\| v_\theta(x_\tau,\tau,c) - (x_1-x_0) \big\|^2 .
\end{equation}
For the \textbf{variance stage}, $x_1$ is the (normalized log-) next variance and
$c=(\vt, r_{t-1}, a_t)$; for the \textbf{return stage}, $x_1$ is the (normalized)
next log-return and $c=(\vt[t{+}1],\vt, r_{t-1}, a_t)$. The velocity field is a
residual MLP (hidden width $256$, $6$ residual blocks, sinusoidal time embedding),
trained with batch size $4096$ for $60$ epochs, cosine-decayed learning rate
$2\!\times\!10^{-4}$, and $10\%$ \emph{action dropout} so that classifier-free
guidance over the regime is available at sampling time. Sampling integrates the
ODE $\dot x = v_\theta(x,\tau,c)$ from $\tau{=}0$ to $1$.

\paragraph{Scheduled sampling.}
Because the teacher is trained one step at a time but deployed in free rollout, it
suffers exposure bias. We optionally train the return stage with
\emph{scheduled sampling} \citep{bengio2015scheduled}: with probability $p$ the
conditioning return $r_{t-1}$ is replaced by the model's own sample rather than the
ground truth. We find $p{=}0.8$ gives the best stylized-fact balance.

\paragraph{Lambert-W heavy-tail variance kernel (LWFM).}
To better capture tail mass we optionally pass returns through a Lambert-W
heavy-tail transform \citep{goerg2015lambert} with parameter $\delta$ before
flow-matching (``LWFM''); inverting the transform at sampling time re-injects the
tails. A small sweep ($\delta\in\{0.03,\dots,0.12\}$) raises kurtosis toward the
real value at a modest pricing cost (\S\ref{sec:results}).

\subsection{Strictly-proper path-distribution fine-tuning (our method)}
\label{sec:pathloss}
The teacher minimizes a \emph{per-step} regression loss; it never optimizes the
\emph{distribution of whole paths} it actually produces under free rollout. Our
method closes this gap. We perform a \emph{differentiable free-running rollout}:
the FM sampler is reparameterized so gradients flow through the $252$-step
generation, producing a batch of generated return paths
$\hat R\in\R^{B\times T}$ that we compare to a batch of real paths $R$ with a sum
of strictly-proper \citep{gneiting2007strictly} path-distribution losses,
\emph{replacing} the WGAN-GP critic of Quant-GAN:
\begin{equation}
\mathcal L
= \lambda_{\mathrm{sig}}\,\mathcal L_{\mathrm{sigMMD}}
+ \lambda_{w_1}\,\mathcal L_{\mathrm{SigW_1}}
+ \lambda_{e}\,\mathcal L_{\mathrm{energy}}
+ \mathcal L_{\mathrm{moment}} ,
\end{equation}
where $\mathcal L_{\mathrm{moment}}$ aggregates low-order moment, terminal-variance,
absolute-return and kurtosis matching plus an anchor regularizer that keeps the
fine-tuned field close to the teacher.

\paragraph{Signature features.} For a standardized return path we form the path
$t\mapsto(t/T,\,\sum_{s\le t}r_s)$ and compute its truncated signature
$\mathbf S(\cdot)$ up to depth $3$ \citep{lyons2007differential}; the signature
linearizes path functionals, so distances in signature space compare whole-path
laws.

\paragraph{Signature-kernel MMD (proper).} With signature features $\phi=\mathbf S$
and an RBF kernel $k$, the (squared) MMD \citep{gretton2012kernel,
salvi2021signature,issa2023nonadversarial} is
\begin{equation}
\mathcal L_{\mathrm{sigMMD}}
= \E\,k(\phi_R,\phi_{R'}) + \E\,k(\phi_{\hat R},\phi_{\hat R'})
- 2\,\E\,k(\phi_R,\phi_{\hat R}),
\end{equation}
a strictly-proper kernel score that is zero iff the path laws coincide.

\paragraph{Sig-W$_1$ (proper).} The expected-signature distance
\citep{ni2021sigwgan,ni2020conditional}
$\mathcal L_{\mathrm{SigW_1}}=\big\|\E[\mathbf S(R)]-\E[\mathbf S(\hat R)]\big\|$
matches all polynomial path functionals up to the truncation depth.

\paragraph{Energy score (proper).} On standardized return paths
\citep{pacchiardi2024probabilistic},
\begin{equation}
\mathcal L_{\mathrm{energy}}
= 2\,\E\,\|R-\hat R\| - \E\,\|\hat R-\hat R'\| - \E\,\|R-R'\|,
\end{equation}
the population energy distance, again strictly proper.

We freeze the variance stage and fine-tune the return stage for $10$ epochs at
learning rate $5\!\times\!10^{-6}$ (batch $512$, rollout ODE in $4$ steps), with
$(\lambda_{\mathrm{sig}},\lambda_{w_1},\lambda_e)=(200,5,10)$. Crucially this is
\emph{non-adversarial}: there is no critic to balance, so training is stable. We
emphasize that the contribution is the \emph{system}---a two-stage FM world model
fine-tuned on a differentiable rollout with a basket of proper path scores---rather
than any single new primitive.

\subsection{Few-step distillation of the world model}
\label{sec:distill}
A 16-step ODE per stage per step is expensive over a 252-step rollout. We distill
the teacher into one-/few-step students.
\textbf{Consistency distillation (CD)} \citep{song2023consistency,song2024improved}
trains a student whose one-step output is invariant along the teacher ODE.
\textbf{Mean-Flow (MF)} \citep{geng2025meanflow} regresses the time-averaged
velocity for one-step sampling.
\textbf{Lagrangian flow-map (FM-map)} \citep{boffi2025flowmaps} learns a two-time
map $X(x,s,t)=x-(t-s)\,G_\theta(x,s,t)$ by enforcing
$\tfrac{d}{dt}X=v_{\mathrm{teacher}}(X,t)$ along the map's own trajectory; this
needs no spatial derivatives and supports $K$-step composition, letting us sweep
the number of function evaluations (NFE) from a single checkpoint.

% =====================================================================
\section{Experiments and Results}
\label{sec:results}

\paragraph{Implementation.} All models train on RTX-5090 GPUs in PyTorch.
Distillation caches data on-device and caps batches to stay compute-bound rather
than I/O-bound. Baselines (below) are fit on CPU.

\paragraph{Baselines.} We compare against (i) \textbf{Quant-GAN}
\citep{wiese2020quant} in its published recipe (TCN generator, Lambert-W output,
WGAN-GP, no moment penalty), reporting both its best and last checkpoints;
(ii) \textbf{GARCH(1,1)-$t$} \citep{bollerslev1986garch,bollerslev1987conditionally},
a Student-$t$ GARCH fit by maximum likelihood on the pooled training returns
($\hat\alpha{=}0.082$, $\hat\beta{=}0.911$, $\hat\nu{=}6.1$); and
(iii) a \textbf{moving-block bootstrap} \citep{kunsch1989jackknife,
politis1994stationary} of real training returns (block length $21$). The latter
two are standard econometric references and need no GPU. Neither is
action-conditional: they cannot be driven into a chosen regime, which is precisely
the controllability the world model adds.

\subsection{Main results}
Table~\ref{tab:main} reports the headline comparison. Three observations.
First, the \textbf{base FM teacher} attains the best \emph{raw} pricing
\emph{among learned generators} ($\best{0.475}$), close to the $0.165$ MC floor and
far ahead of GARCH-$t$ ($1.089$) and Quant-GAN ($5.4$--$5.7$); its near-equal
\raw{}/\cal{} values ($0.475/0.583$) show it is genuinely self-consistent rather
than relying on calibration. The \textbf{block-bootstrap} prices even better
($0.269$ raw, kurtosis $4.58$), but this is expected and not a fair comparison: it
simply replays real return blocks, so it is a non-parametric \emph{skyline} on
in-sample marginals---it cannot be driven to a chosen regime and cannot generate
genuinely novel scenarios, which is precisely the controllable, extrapolative
capability the world model provides. Second, \textbf{scheduled sampling}
(FM\,+\,SS) gives the best stylized-fact balance: kurtosis $4.41$ against the real
$4.60$, with a strong calibrated price of $0.348$. Third, our
\textbf{strictly-proper fine-tuning} delivers the best \emph{honest} calibrated
pricing ($\best{0.180}$), improving over the adversarial critic (cf.\
Table~\ref{tab:ablation}). The classical \textbf{GARCH-$t$} baseline prices
surprisingly well raw ($1.089$) because it nails the unconditional variance, but
its kurtosis explodes to $19.2$ and it has no regime control---illustrating that
good pricing and good path realism are distinct axes.

\begin{table}[t]
\centering
\caption{Main comparison on the three-regime Heston task under free-running
rollout. \raw{}/\cal{} = uncalibrated / moment-calibrated option-pricing RMSE
against the MC oracle (lower is better; finite-sample floor $0.165$). Kurtosis
target $4.60$. $^{\dagger}$\cal{} below the floor: calibration artifact, not
generative quality. $^{\ddagger}$non-parametric resampling \emph{skyline}: reuses
real returns, so its near-floor marginals are expected; it is \emph{not}
controllable (cannot be driven to a chosen regime) and cannot extrapolate to
unseen scenarios, so it is a reference rather than a competing generator.
$^{\ast}$best among \emph{learned} generators. Best learned value per column in
\best{bold}.}
\label{tab:main}
\small
\begin{tabular}{l l c c c}
\toprule
& Model & \raw{} $\downarrow$ & \cal{} $\downarrow$ & Kurt.\ ($\to 4.60$) \\
\midrule
\multirow{2}{*}{\shortstack[l]{\emph{Reference}\\\emph{(non-gen.)}}}
 & Real test set (vs.\ oracle) & 0.165 & --- & 4.60 \\
 & Block bootstrap$^{\ddagger}$ (data replay) \citep{politis1994stationary} & 0.269 & 0.244 & 4.58 \\
\midrule
\multirow{3}{*}{\emph{Baselines}}
 & Quant-GAN (paper, best) \citep{wiese2020quant} & 5.650 & 1.607 & 10.21 \\
 & Quant-GAN (paper, last) \citep{wiese2020quant} & 5.429 & 0.674 & 6.39 \\
 & GARCH(1,1)-$t$ \citep{bollerslev1986garch}      & 1.089 & 0.482 & 19.17 \\
\midrule
\multirow{4}{*}{\emph{Teachers}}
 & FM world model (base)                 & \best{0.475}$^{\ast}$ & 0.583 & 3.88 \\
 & \quad + LWFM tail kernel ($\delta{=}0.05$) & 0.882 & --- & 4.33 \\
 & \quad + scheduled sampling ($p{=}0.8$) & 0.872 & 0.348 & \best{4.41} \\
 & Strong-recipe teacher (256/6, bs 8192) & 1.057 & 0.582 & 4.08 \\
\midrule
\multirow{3}{*}{\shortstack[l]{\emph{Our method}\\(path-loss\\fine-tune)}}
 & on strong teacher (base recipe)        & 2.595 & 0.232 & 3.66 \\
 & \quad + heavy moment/anchor (\textbf{best cal}) & 2.595 & \best{0.180} & 3.63 \\
 & \quad long-train variant               & 1.483 & 0.283 & 3.68 \\
\midrule
\multirow{2}{*}{\emph{Distilled}}
 & Consistency student of our method      & 1.875 & 0.170$^{\dagger}$ & 3.11 \\
 & \quad (low-lr, strongest calibration)  & 2.328 & 0.118$^{\dagger}$ & 2.85 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ablation: each proper score alone is insufficient}
\label{sec:ablation-path}
Table~\ref{tab:ablation} isolates the components of the path-loss
(\S\ref{sec:pathloss}), all on the same LWFM teacher and rollout. Every
\emph{single} proper score, used alone, is poor: the signature scores
($\mathcal L_{\mathrm{sigMMD}}$, $\mathcal L_{\mathrm{SigW_1}}$) destabilize
calibration (\cal{} $>10$), the energy score and moment-only control plateau
around \cal{} $1.5$--$2.9$. The \emph{adversarial} WGAN-GP critic (the Quant-GAN
objective ported onto our rollout) reaches \cal{} $0.420$. Only the
\emph{combined} strictly-proper basket reaches \cal{} $0.362$ on this teacher
(and $0.180$ on the strong teacher, Table~\ref{tab:main}), confirming that the
contribution is the \emph{integration}: the scores are complementary, and together
they beat the adversarial critic without its instability.

\begin{table}[t]
\centering
\caption{Component ablation of the path-distribution loss (same teacher/rollout).
Each single strictly-proper score is insufficient; the combination beats the
adversarial WGAN-GP critic.}
\label{tab:ablation}
\small
\begin{tabular}{l c c c}
\toprule
Path-loss configuration & \raw{} $\downarrow$ & \cal{} $\downarrow$ & Kurt.\ ($\to 4.60$) \\
\midrule
Moment-only (control)                       & 2.566 & 2.904 & 3.49 \\
sig-MMD only                                & 5.510 & 15.08 & 4.78 \\
Sig-W$_1$ only                              & 4.971 & 11.79 & 4.59 \\
Energy only                                 & 1.814 & 1.555 & 3.39 \\
WGAN-GP critic (Quant-GAN objective)        & 1.746 & 0.420 & 4.21 \\
\textbf{Combined (sig-MMD + Sig-W$_1$ + energy + moment)} & 1.717 & \best{0.362} & 4.22 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Distillation study}
\label{sec:results-distill}
Table~\ref{tab:distill} reports one-step students. \textbf{Consistency
distillation} transfers across teachers: it preserves the balanced student's shape
(B1$\to$CD: kurtosis $4.27$, \raw{} $1.28$) and inherits the calibrated-pricing
strength of our method (combined$\to$CD: \cal{} $0.170$). In contrast
\textbf{Mean-Flow fails} on this task regardless of teacher---the FM-teacher MF
student degrades to \raw{} $9.08$, and the our-method MF student blows up
(\cal{} $37.6$, kurtosis collapses to $2.49$). We attribute this to MF's
time-averaged-velocity target being a poor fit to the sharply state-dependent,
heavy-tailed return conditional under \emph{compounded} free rollout: small
one-step biases accumulate over $252$ steps. The strongest calibrated numbers
(\cal{} $0.118$--$0.170$) come from distilling our method, but the sub-floor ones
carry the calibration-artifact caveat and have poor \raw{}/kurtosis---they are
\emph{pricing accelerators}, not better generators.

\begin{table}[t]
\centering
\caption{One-step distillation. Consistency distillation (CD) transfers under free
rollout; Mean-Flow (MF) fails. $^{\dagger}$ sub-floor \cal{}: calibration artifact.}
\label{tab:distill}
\small
\begin{tabular}{l l c c c}
\toprule
Teacher & Student & \raw{} $\downarrow$ & \cal{} $\downarrow$ & Kurt.\ ($\to 4.60$) \\
\midrule
FM base        & CD & 4.810 & 0.537 & 4.48 \\
FM base        & MF (fail) & 9.084 & 1.442 & 3.52 \\
FM\,+\,SS (B1) & CD & 1.282 & 0.371 & 4.27 \\
FM\,+\,SS (B1) & CD (low-lr)  & \best{1.005} & 0.444 & 4.27 \\
Our method     & CD & 1.875 & 0.170$^{\dagger}$ & 3.11 \\
Our method     & CD (low-lr)  & 2.328 & 0.118$^{\dagger}$ & 2.85 \\
Our method     & MF (fail) & 7.536 & 37.62 & 2.49 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Few-step flow-map: the NFE knee}
\label{sec:results-nfe}
Table~\ref{tab:nfe} sweeps the Lagrangian flow-map student over
$K\in\{1,2,4,8\}$ steps. The optimum depends on the metric: \emph{raw} pricing is
best at $K{=}2$ ($1.82$) while \emph{calibrated} pricing is best at $K{=}1$
($0.313$). Adding steps does \emph{not} monotonically help---kurtosis inflates from
$4.86$ ($K{=}1$) to $8.10$ ($K{=}8$) as composition amplifies tail errors. The
flow-map student is competitive at very low NFE but does not match consistency
distillation, and a warm-start sign audit (negating the FM output layer) did not
fix it, indicating the weakness is a rollout/path-distribution mismatch rather
than an initialization artifact.

\begin{table}[t]
\centering
\caption{Few-step Lagrangian flow-map student: NFE ($K$) sweep from one checkpoint.
The quality knee differs by metric (\raw{}: $K{=}2$; \cal{}: $K{=}1$); more steps
inflate the tails.}
\label{tab:nfe}
\small
\begin{tabular}{l c c c c}
\toprule
$K$ (NFE) & 1 & 2 & 4 & 8 \\
\midrule
\raw{} $\downarrow$  & 2.736 & \best{1.819} & 3.090 & 3.437 \\
\cal{} $\downarrow$  & \best{0.313} & 2.355 & 3.745 & 3.857 \\
Kurtosis ($\to 4.60$) & 4.86 & 5.53 & 7.85 & 8.10 \\
\bottomrule
\end{tabular}
\end{table}

% =====================================================================
\section{Comparative Analysis and Discussion}
\label{sec:discussion}

\paragraph{The raw$\leftrightarrow$calibrated trade-off.}
Across Tables~\ref{tab:main}--\ref{tab:distill}, every procedure that lowers
\emph{calibrated} pricing or moves kurtosis toward the target costs \emph{raw}
self-consistency, and vice versa. The base FM teacher sits at one extreme
(\raw{} $0.475$, near-equal \cal{}); our path-loss and the CD students sit at the
other (\cal{} as low as $0.118$ but \raw{} $>2$ and kurtosis $<3$). Scheduled
sampling is the most attractive interior point. This trade-off is the central
empirical message: optimizing the two pricing-dominant moments (which calibration
or a terminal-variance-heavy loss does) trades away path-level realism.

\paragraph{Calibration artifacts.}
Because option prices are dominated by terminal variance, pinning the pooled
return mean and standard deviation can drive \cal{} below the $0.165$ finite-sample
floor. Our most aggressive consistency student reaches \cal{} $0.118$ while its
\raw{} error is $2.33$ and its kurtosis is only $2.85$; another reaches
\cal{} $0.170$ with kurtosis $3.11$. These are \emph{not} better
generators---they are artifacts of injecting two data statistics, as their poor
\raw{} and kurtosis make plain. We therefore advocate reporting \raw{} alongside
\cal{} and treating any sub-floor \cal{} as diagnostic only. We believe this
distinction is under-appreciated in the financial-generation literature, where
single calibrated pricing numbers are common.

\paragraph{Why Mean-Flow and flow-map struggle here.}
Both target a (time-)averaged transport in one or few steps. Under a 252-step
\emph{compounded} rollout, the per-step return conditional is sharp and
heavy-tailed and small averaging biases accumulate multiplicatively in the price
path; consistency distillation, which matches the ODE trajectory pointwise, is more
robust. This is a concrete, quantitative negative result for one-step generative
modeling of long autoregressive simulators.

\paragraph{Strictly-proper vs.\ adversarial.}
Our combined path-loss beats the WGAN-GP critic (\cal{} $0.362$ vs.\ $0.420$ on the
same teacher; $0.180$ best overall) while being non-adversarial and stable. This
supports the broader thesis \citep{issa2023nonadversarial,pacchiardi2024probabilistic}
that strictly-proper scoring rules are a practical, stable alternative to
discriminators for path generation---here, inside a world-model fine-tune.

\paragraph{Limitations.}
Our method improves calibrated pricing but not raw self-consistency; post-hoc
fine-tuning cannot have both. We conjecture the proper-scoring path loss should be
\emph{baked into teacher training and checkpoint selection} rather than applied
post-hoc---a direction we leave to future work. We also study a synthetic (if
realistic) market; transfer to real microstructure data is open.

% =====================================================================
\section{Conclusion and Future Work}
\label{sec:conclusion}

We presented an action-conditional, two-stage Flow-Matching \emph{world model} for
regime-switching financial markets and two contributions on top of it: a
\emph{strictly-proper path-distribution fine-tuning} that replaces the adversarial
Quant-GAN critic and yields the best honest calibrated option-pricing ($0.180$),
and a \emph{systematic distillation study} showing that consistency distillation
transfers under free rollout while Mean-Flow and flow-map do not. Our honest
evaluation---free-running rollout, pricing against an MC oracle, and a careful
\raw{}/\cal{} split with an explicit floor---exposes a robust
raw$\leftrightarrow$calibrated trade-off and a class of calibration artifacts that
single calibrated numbers hide. The base FM teacher remains the raw
self-consistency champion ($0.475$); scheduled sampling gives the best stylized-fact
balance (kurtosis $4.41$); our method gives the best calibrated pricing; and
consistency distillation makes the simulator cheap without losing fidelity.

\paragraph{Future work.} (i) Integrate the proper-scoring path loss into the
teacher's native training objective and checkpoint selection, targeting low
\raw{} \emph{and} \cal{} simultaneously. (ii) A multistep consistency sampler
(currently only flow-map/MF support the NFE sweep). (iii) Architectural changes
(e.g.\ attention over the rollout) to attack exposure bias directly. (iv) Transfer
to real limit-order-book and equity data with the same world-model and evaluation
protocol.

\paragraph{Reproducibility.} We release all code: the path-loss fine-tune
(\texttt{pathwise\_teacher\_combined.py}), the three distillation recipes
(\texttt{distill\_\{consistency,mean\_flow,flow\_map\}.py}), the few-step sampler
(\texttt{rollout\_fewstep.py}), and the evaluation pipeline
(\texttt{rollout.py}, \texttt{evaluate\_rollout.py}). Data are regenerated from
\texttt{generate\_heston\_data.py} (seed $20260530$) and
\texttt{generate\_mc\_oracle.py}.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
